Modeling the Linux page cache for accurate simulation of data-intensive applications

### 2021-05-20 paper_io_sim.tex
A first pass, not the last one

        %To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}
        Page cache offsets the cost of disk I/O by enabling I/O to occur directly from memory.
        When a file is first loaded into memory, the file is read from disk and loaded into the page cache as
        a series of pages. Subsequent reads to any of the file pages located in memory will result in
        a \textit{cache hit}, meaning the I/O can occur directly from without disk involvement.
        Any accessed page not loaded in memory results in a \textit{cache miss}, resulting in
        the page being read directly from disk.

@@

        Cache eviction and flushing strategies are integral to proper page cache functioning.
        Whenever space in memory becomes limited, either as a result of application memory
        or page cache use, page cache data may be evicted. Only data that
        has been persisted to storage (clean pages) can be flagged for eviction and removed from
        memory. Written data that has not yet been persisted to disk (dirty data) must first
        be copied (flushed) to storage prior to eviction. When sufficient memory is
        being occupied, the flushing process is synchronous. However, even when
        there is sufficient available memory, written data will be flushed to disk
        at a predefined interval through a process known as \textit{periodical flushing}.
        Periodical flushing only flushes ``expired" dirty pages, that is, pages that
        have remained dirty in the
        page cache longer than an expiration time configured in the kernel.
        Different cache eviction algorithms have also been proposed
        \cite{owda2014comparison}.

@@

        Although the Linux page cache has a large impact on I/O
        performance, and thus on the execution of data-intensive
        applications, its simulation is rarely considered in the above
        frameworks.  Most frameworks merely simulate I/O operations
        based on storage bandwidths and capacities.  The SIMCAN
        framework does models page caching by storing data accessed on
        disk in a block cache~\cite{nunez2012simcan}.  Page cache is
        also modeled in iCanCloud through a component that manages
        memory accesses and cached data~\cite{nunez2012icancloud}.
        However, the scalability of these simulators is limited
        as they rely on microscopic simulation models.  Besides, none
        of these simulators provide any writeback cache simulator nor
        cache eviction policies through LRU lists.  Although cache
        replacement policies are applied in~\cite{xu2018saving} to
        simulate in-memory caching, this simulator is specific to
        energy consumption of multi-tier heterogeneous networks.

@@

    \begin{figure}
           \centering
           \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
           \caption{Overview of the page cache simulator.
           Applications send file read or write requests to the
           I/O Controller that orchestrates flushing, eviction, cache
           and disk accesses with the Memory Manager. Concurrent accesses to storage
           devices (memory and disk) are simulated using existing models.}
           \label{fig:interaction}
    \end{figure}

@@

    The Memory Manager simulates two parallel threads: the main one
    implements flushing, eviction, and cached I/Os synchronously, whereas
    the second one, which operates in the background, periodically searches for
    expired dirty data in LRU lists and flushes this data to disk. We
    use existing storage simulation models~\cite{lebre2015} to simulate disk and
    memory, characterized by their storage capacity, read and write
    bandwidths, and latency. These models account for
    bandwidth sharing between concurrent memory or disk accesses.

@@

    Our simulation model supports chunk-by-chunk file accesses
    with a user-defined chunk size. However, for simplicity, we assume that file pages are
    accessed in a round-robin fashion rather than fully randomly.
    Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
    before data from the active list
    (data reads occur from left to right in Figure~\ref{fig:read_order}).
    When a chunk of \emph{uncached} data is read, a new clean block is created
    and appended to the inactive list.
    When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
    If these blocks are clean, we merge them together, update the access time and size of the resulting block,
    and append it to the active list.
    If the blocks are dirty, we move them independently to the active list, to preserve their entry time.
    Because the chunk and block sizes may be different, there are situations
    where a block is not entirely read.
    In this case, the block is split into two smaller blocks and one of them is re-accessed.
    % From Tristan: not sure where to put this nor if it's necessary:
    % , in which chunks are read/written
    % until file is entirely read/written.
    \begin{figure}
           \centering
           \includegraphics[width=\columnwidth]{figures/read_order.pdf}
           \caption{File data read order. Data is read from left to right: uncached data
           is read first, followed by data from the inactive list, and finally data from the active list. }
           \label{fig:read_order}
    \end{figure}

@@

    The above model describes page cache in writeback
    mode. Our model also includes a write function in writethrough mode,
    which simply simulates a disk write with the amount of data passed in,
    performs cache eviction if needed and adds the written data to the cache.

@@

            This prototype does not simulate  bandwidth sharing and thus does not support
            concurrency: it is limited to single-threaded applications running on systems
            with a single-core CPU. We used this prototype for a first validation of our simulation
            model against a real sequential application running on a real system.
            The Python 3.7 source code is available at
            \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}.

@@

        Experiments are run on a dedicated cluster at
        Concordia University, with one login node, 9 compute nodes, and 4
        storage nodes connected with a 25 Gbps network. Each
        compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
        @ 2.10GHz, 250~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
        file system, 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
        CentOS~8.1 and NFS version 4. We use the \texttt{atop}
        and \texttt{collectl} tools to monitor and collect memory status
        and disk throughput. We clear the page
        cache before each application run to ensure comparable
        conditions.

        \begin{table}[b]
            \centering
            \begin{tabularx}{0.8\columnwidth}{c>{\centering\arraybackslash}X}
            \toprule
                Input size (GB)  & CPU time (s)\\
            \midrule
                3      & 4.4 \\
                20  & 28 \\
                50  & 75 \\
                75  & 110 \\
                100  & 155 \\
            \bottomrule
            \end{tabularx}
            \caption{Synthetic application parameters}
            \label{table:cputime}
            \end{table}
        The synthetic application, implemented in C, consists of three single-core,
        sequential tasks where each task reads the file produced by the
        previous task, increments every byte of this file to emulate real
        processing, and writes the resulting data to disk. Files are
        numbered by ascending access times (File 1 is the file read by Task 1, etc).
         The anonymous memory used by the application
        is released after each task, which we also simulate in the Python
        prototype and in \wrench-cache. As our focus is on I/O rather than compute, we measure
        application task CPU times on a cluster node
        (Table~\ref{table:cputime}), and use these durations in our
        simulations. For the Python prototype, we inject CPU times
        directly into the simulation. For \wrench and \wrench-cache, we
        determined the corresponding number of flops on a 1~Gflops CPU
        and used these values in the simulation. The simulated
        platform and application are available at
        commit \href{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}{ec6b43561b}.

        We used the synthetic application in three experiments. In the
        first one (\textit{Exp~1}), we run \emph{a single instance} of
        the application on a single cluster node, with different input file
        sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed to
        the same local disk.
        In the second experiment (\textit{Exp~2}), we run
        \emph{concurrent} instances of the application on a single node,
        all application instances operating on different files of size 3~GB
        stored in the same local disk. We vary the number of concurrent
        application instances from 1 to 32 since cluster nodes have 32 CPU
        cores.
        In the third experiment (\textit{Exp~3}), we use the same
        configuration as the previous one, albeit reading and writing
        on a 50-GiB NFS-mounted partition of a 450-GiB remote disk of
        another compute node. As is commonly configured in HPC
        environments to avoid data loss, there is no client write cache
        and the server cache is configured as writethrough instead of
        writeback. NFS client and server read caches are enabled. 
        Therefore, all the writes happen at disk bandwidth, but
        reads can benefit from cache hits.

        The real application is a workflow of the Nighres
        toolbox~\cite{huntenburg2018nighres}, which implements cortical
        reconstruction from brain images in four steps: skull stripping,
        tissue classification, region extraction, and cortical
        reconstruction. Each step reads files produced by the previous step,
        and writes files that are or are not read by the subsequent step.
        More information on this application is available in the Nighres
        documentation at
        \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}.
        The application is implemented as a Python script that calls Java
        image-processing routines. We use Python 3.6, Java 8, and Nighres
        1.3.0. We patched the application to remove lazy data loading and
        data compression, which made CPU time difficult to separate from
        I/O time, and to capture task CPU times to inject in the
        simulation. The patched code is available at
        \url{https://github.com/dohoangdzung/nighres}.
        \begin{table}[t]
            \centering
            \begin{tabular}{lccc}
            \toprule
                \multicolumn{1}{c}{Workflow step}& Input size       & Output size      & CPU time\\
                                       & (MB)             & (MB)             & (s)\\
            \midrule
               Skull stripping         &  295             & 393               & 137 \\
               Tissue classification   &  197              & 1376              & 614 \\
               Region extraction       &  1376             & 885              & 76 \\
               Cortical reconstruction &  393              & 786              & 272\\
            \bottomrule
            \end{tabular} 
            \caption{Nighres application parameters}
            \label{table:nighres_stats}
            \end{table}
            \begin{table}[b]
                \centering
                \begin{tabularx}{\columnwidth}{ll
                >{\centering\arraybackslash}X
                >{\centering\arraybackslash}X
                >{\centering\arraybackslash}X}
                \toprule
                    \multicolumn{2}{c}{Bandwidths}  & Cluster (real) & Python prototype & \wrench simulator\\
                \midrule
                \multirow{2}{*}{Memory}      & read  & 6860 & 4812 & 4812\\
                                             & write & 2764 & 4812 & 4812\\
                \multirow{2}{*}{Local disk}  & read  & 510  & 465  & 465\\
                                             & write & 420  & 465  & 465\\
                \multirow{2}{*}{Remote disk} & read  & 515  & -    & 445\\
                                             & write & 375  & -    & 445\\
                \multicolumn{2}{l}{Network}  & 3000  & -    & 3000\\
                \bottomrule
                \end{tabularx}
                \caption{Bandwidth benchmarks (MBps) and simulator configurations.
                The bandwidths used in the simulations were the average of the measured read and write bandwidths.
                Network accesses were not simulated in the Python prototype.}
                \label{table:benchmark}
                \end{table}

        We use the real application in the fourth experiment
        (\textit{Exp~4}), which runs on a single cluster node 
        using a single local disk, to process data from
        participant 0027430 in the dataset of the Max Planck Institute for
        Human Cognitive and Brain Sciences available at
        \url{http://dx.doi.org/10.15387/fcp_indi.corr.mpg1}, leading to the
        parameters in Table~\ref{table:nighres_stats}.

        To calibrate the simulators, we benchmark the
        memory, local disk, remote disk (NFS), and network bandwidths
        (Table~\ref{table:benchmark}). Since \simgrid, and thus \wrench, currently only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.

@@

    \begin{figure*}
        \centering
        \begin{subfigure}{\linewidth}
            \centering
               \includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
               \vspace*{-0.7cm}
               \caption{Relative simulation errors (absolute values)}
               \vspace*{0.5cm}
               \label{fig:single_error}
            \end{subfigure}
        \begin{subfigure}{\linewidth}
            \centering
            %    Gray shades represent task phases (read, compute and write).
            %    Lines represent memory usage along pipeline execution time.}
               \includegraphics[width=\linewidth]{result/single/figures/single_memprof.pdf}
               \vspace*{-0.7cm}
               \caption{Memory profiles}
               \vspace*{0.5cm}
               \label{fig:single_memprof}
        \end{subfigure}
        \begin{subfigure}{\linewidth}
            \centering
               \includegraphics[width=1.05\linewidth]{result/single/figures/cached_files.pdf}
               \caption{Cache contents \emph{after} application I/O operations}
            %    \textcolor{red}{Update real results of 20GB}}
               \label{fig:single_cache}
        \end{subfigure}
        \caption{Single-threaded results (\textit{Exp 1})}
        \end{figure*}

@@

        \wrench simulation errors were substantially lower with 100~GB
        files than with 20~GB files, due to the fact that part of the
        100~GB file needs to be read and written to disk, the only storage
        device in \wrench, as it does not fit in cache. Conversely,
        simulation errors of the Python prototype and \wrench-cache are higher with
        100~GB files than with 20~GB files, due to idiosyncrasies in the kernel
        flushing and eviction strategies that could not be easily modeled.

        Simulated memory profiles are highly consistent with the real ones
        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory profiles almost exactly match the
        real ones, although dirty data seems to be flushing faster in real
        life than in simulation, a behavior also
        observed with 100~GB files. With 100~GB files, used memory reaches
        total memory during the first write, triggering dirty data
        flushing, and drops back to cached memory when application tasks
        release anonymous memory. Simulated cached memory is highly
        consistent with real values, except toward the end of Read 3 where
        it slightly increases in simulation but not in reality. This
        occurs because after Write 2 File 3 is only partially
        cached in simulation whereas it is entirely cached in the real
        system. In all cases, dirty data remains under the dirty ratio as
        expected. The Python prototype and \wrench-cache exhibit nearly
        identical memory profiles, which reinforces the confidence in our
        implementations.

        The content of the simulated memory cache is also highly
        consistent with reality (Figure~\ref{fig:single_cache}). With 20~GB
        files, the simulated cache content exactly match reality, since
        all files fit in page cache. With 100~GB files, a slight
        discrepancy is observed after Write 2, which explains the
        simulation error previously mentioned in Read 3. Indeed, in the real
        execution File 3 is entirely cached after Write 2, whereas
        in the simulated execution only part of it is cached. This is because
        the Linux kernel tends to not evict pages that
        belong to files being currently written (File 3 in this case),
        which we cannot easily reproduce in our model.

            \begin{figure*}
                \begin{subfigure}{\linewidth}
                    \centering
                    \includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
                \end{subfigure}
                \caption{Concurrent results with 3~GB files (\textit{Exp 2})}
                \label{fig:multi_local}
            \end{figure*}
        \subsection{Concurrent applications (Exp 2)}

@@

            \begin{figure}[b]
                \begin{subfigure}{0.95\linewidth}
                    \centering
                    \includegraphics[width=\linewidth]{result/nighres/figures/nighres_errors.pdf}
                \end{subfigure}
                \caption{Real application results (\textit{Exp 4})}
                \label{fig:nighres}
                \end{figure}

@@

            \begin{figure*}
                \begin{subfigure}{\linewidth}
                    \centering
                    \includegraphics[width=\linewidth]{result/multi/figures/multi_nfs.pdf}
                \end{subfigure}
                \caption{NFS results with 3~GB files (\textit{Exp 3})}
                \label{fig:multi_nfs}
                \end{figure*}

            \begin{figure}
                \begin{subfigure}{\columnwidth}
                    \centering
                    \includegraphics[width=\linewidth]{result/multi/figures/simulation_time.pdf}
                \end{subfigure}
                \caption{Simulation time comparison. \wrench-cache scales
                linearly with the number of concurrent applications, albeit
                with a higher overhead than \wrench.}
                \label{fig:multi_time}
                \end{figure}

@@

        As is the case for \wrench, simulation time with \wrench-cache scales
        linearly with the number of concurrent applications
        (Figure~\ref{fig:multi_time}, p \textless $10^{-24}$). However, the page
        cache model substantially increases simulation time by
        application, as can be seen by comparing regression slopes in
        Figure~\ref{fig:multi_time}. Interestingly, \wrench-cache is faster with 
        NFS I/Os than with local I/Os, most likely due to the use of writethrough
        cache in NFS, which bypasses flushing operations.

@@

\bibliographystyle{IEEEtran}
\bibliography{citation}

>>>>

        %To improve performance of disk accesses, the page cache was introduced to the Linux kernel. %It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}
        Page cache offsets the cost of disk I/O by enabling I/O to occur directly from memory.
        When a file is first loaded into memory, the file is read from disk and loaded into the page cache as
        a series of pages. Subsequent reads to any of the file pages located in memory will result in
        a cache hit, meaning the I/O can occur directly from without disk involvement.
        Any accessed page not loaded in memory results in a cache miss, resulting in
        the page being read directly from disk.

@@

        Cache eviction and flushing strategies are integral to proper page cache functioning.
        Whenever space in memory becomes limited, either as a result of application memory
        or page cache use, page cache data may be evicted. Only data that
        has been persisted to storage (clean pages) can be flagged for eviction and removed from
        memory. Written data that has not yet been persisted to disk (dirty data) must first
        be copied (flushed) to storage prior to eviction. When sufficient memory is
        being occupied, the flushing process is synchronous. However, even when
        there is sufficient available memory, written data will be flushed to disk
        at a predefined interval through a process known as periodical flushing.
        Periodical flushing only flushes ``expired" dirty pages, that is, pages that
        have remained dirty in the
        page cache longer than an expiration time configured in the kernel.
        Different cache eviction algorithms have also been proposed
        \cite{owda2014comparison}.

@@

        Although the Linux page cache has a large impact on I/O
        performance, and thus on the execution of data-intensive
        applications, its simulation is rarely considered in the above
        frameworks.  Most frameworks merely simulate I/O operations
        based on storage bandwidths and capacities.  The SIMCAN
        framework does model page caching by storing data accessed on
        disk in a block cache~\cite{nunez2012simcan}.  Page cache is
        also modeled in iCanCloud through a component that manages
        memory accesses and cached data~\cite{nunez2012icancloud}.
        However, the scalability of these simulators is limited
        as they rely on microscopic simulation models.  Besides, none
        of these simulators provide any writeback cache simulator nor
        cache eviction policies through LRU lists.  Although cache
        replacement policies are applied in~\cite{xu2018saving} to
        simulate in-memory caching, this simulator is specific to
        energy consumption of multi-tier heterogeneous networks.

@@

    \begin{figure}
           \centering
           \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
           \caption{Overview of the page cache simulator.
           Applications send file read or write requests to the
           I/O Controller that orchestrates flushing, eviction, cache
           and disk accesses with the Memory Manager. Concurrent accesses to storage
           devices (memory and disk) are simulated using the existing SimGrid models.}
           \label{fig:interaction}
    \end{figure}

@@

    The Memory Manager simulates two parallel threads: the main one
    implements flushing, eviction, and cached I/Os synchronously, whereas
    the second one, which operates in the background, periodically searches for
    expired dirty data in LRU lists and flushes this data to disk. We
    rely on the existing SimGrid storage simulation models~\cite{lebre2015} to simulate disk and
    memory, characterized by their storage capacity, read and write
    bandwidths, and latency. These models account for
    bandwidth sharing between concurrent memory or disk accesses.

@@

    Our simulation model supports chunk-by-chunk file accesses
    with a user-defined chunk size. However, since file access pattern models are currently not available in SimGrid, we assume that file pages are
    accessed in a round-robin fashion rather than fully randomly.
    Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
    before data from the active list. Such an access pattern is commonly encountered in scientific data processing, where 
    applications often read data files entirely and possibly multiple times. 
    When a chunk of \emph{uncached} data is read, a new clean block is created
    and appended to the inactive list.
    When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
    If these blocks are clean, we merge them together, update the access time and size of the resulting block,
    and append it to the active list.
    If the blocks are dirty, we move them independently to the active list, to preserve their entry time.
    Because the chunk and block sizes may be different, there are situations
    where a block is not entirely read.
    In this case, the block is split into two smaller blocks and one of them is re-accessed.
    % From Tristan: not sure where to put this nor if it's necessary:
    % , in which chunks are read/written
    % until file is entirely read/written.
    \begin{figure}
           \centering
           \includegraphics[width=\columnwidth]{figures/read_order.pdf}
           \caption{File data read order. Data is read from left to right: uncached data
           is read first, followed by data from the inactive list, and finally data from the active list. }
           \label{fig:read_order}
    \end{figure}

@@

    The above model describes page cache in writeback
    mode. Our model also includes a write function in writethrough mode,
    which simply simulates a disk write with the amount of data passed in,
    performs cache eviction if needed and adds the written data to the
    cache. The read function remains unchanged between the writeback and
    writethrough modes.

@@

            This prototype does not simulate  bandwidth sharing and thus
            does not support concurrency: it is limited to single-threaded
            applications running on systems with a single-core CPU. We used
            this prototype for a first validation of our simulation model
            against a real sequential application running on a real system,
            and as a way to detect potential bugs in our implementation.
            The Python 3.7 source code is available at
            \url{https://github.com/big-data-lab-team/paper-io-simulation/tree/master/exp/pysim}.

@@

        Experiments were run on a dedicated cluster at
        Concordia University, with one login node, 9 compute nodes, and 4
        storage nodes connected with a 25 Gbps network. Each
        compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
        @ 2.10GHz, 250~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
        file system, 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
        CentOS~8.1 and NFS version 4. We used the \texttt{atop}
        and \texttt{collectl} tools to monitor and collect memory status
        and disk throughput. We cleared the page
        cache before each application run to ensure comparable
        conditions.

        \begin{table}[b]
            \centering
            \begin{tabularx}{0.8\columnwidth}{c>{\centering\arraybackslash}X}
            \toprule
                Input size (GB)  & CPU time (s)\\
            \midrule
                3      & 4.4 \\
                20  & 28 \\
                50  & 75 \\
                75  & 110 \\
                100  & 155 \\
            \bottomrule
            \end{tabularx}
            \caption{Synthetic application parameters}
            \label{table:cputime}
            \end{table}
        The synthetic application, implemented in C, consists of three single-core,
        sequential tasks where each task reads the file produced by the
        previous task, increments every byte of this file to emulate real
        processing, and writes the resulting data to disk. Files are
        numbered by ascending access times (File 1 is the file read by Task 1, etc).
         The anonymous memory used by the application
        is released after each task, which we also simulate in the Python
        prototype and in \wrench-cache. As our focus is on I/O rather than compute, we measured
        application task CPU times on a cluster node
        (Table~\ref{table:cputime}), and used these durations in our
        simulations. For the Python prototype, we injected CPU times
        directly into the simulation. For \wrench and \wrench-cache, we
        determined the corresponding number of flops on a 1~Gflops CPU
        and used these values in the simulation. The simulated
        platform and application are available at
        commit \href{https://github.com/wrench-project/wrench/tree/ec6b43561b95977002258c0fe37a4ecad8f1d33f/examples/basic-examples/io-pagecache}{ec6b43561b}.

        We used the synthetic application in three experiments. In the
        first one (\textit{Exp~1}), we run \emph{a single instance} of
        the application on a single cluster node, with different input file
        sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed to
        the same local disk.
        In the second experiment (\textit{Exp~2}), we run
        \emph{concurrent} instances of the application on a single node,
        all application instances operating on different files of size 3~GB
        stored in the same local disk. We varied the number of concurrent
        application instances from 1 to 32 since cluster nodes have 32 CPU
        cores.
        In the third experiment (\textit{Exp~3}), we use the same
        configuration as the previous one, albeit reading and writing
        on a 50-GiB NFS-mounted partition of a 450-GiB remote disk of
        another compute node. NFS client and server read caches were enabled, however, as is commonly configured in HPC
        environments to avoid data loss, there was no client write cache
        and the server cache was configured as writethrough instead of
        writeback.
        Therefore, all the writes happened at disk bandwidth, but
        reads could benefit from cache hits.

        The real application is a workflow of the Nighres
        toolbox~\cite{huntenburg2018nighres}, which implements cortical
        reconstruction from brain images in four steps: skull stripping,
        tissue classification, region extraction, and cortical
        reconstruction. Each step reads files produced by the previous step,
        and writes files that are or are not read by the subsequent step.
        More information on this application is available in the Nighres
        documentation at
        \url{https://nighres.readthedocs.io/en/latest/auto_examples/example_02_cortical_depth_estimation.html}.
        The application is implemented as a Python script that calls Java
        image-processing routines. We used Python 3.6, Java 8, and Nighres
        1.3.0. We patched the application to remove lazy data loading and
        data compression, which made CPU time difficult to separate from
        I/O time, and to capture task CPU times to inject in the
        simulation. The patched code is available at
        \url{https://github.com/dohoangdzung/nighres}.
        \begin{table}[t]
            \centering
            \begin{tabular}{lccc}
            \toprule
                \multicolumn{1}{c}{Workflow step}& Input size       & Output size      & CPU time\\
                                       & (MB)             & (MB)             & (s)\\
            \midrule
               Skull stripping         &  295             & 393               & 137 \\
               Tissue classification   &  197              & 1376              & 614 \\
               Region extraction       &  1376             & 885              & 76 \\
               Cortical reconstruction &  393              & 786              & 272\\
            \bottomrule
            \end{tabular} 
            \caption{Nighres application parameters}
            \label{table:nighres_stats}
            \end{table}
            \begin{table}[b]
                \centering
                \begin{tabularx}{\columnwidth}{ll
                >{\centering\arraybackslash}X
                >{\centering\arraybackslash}X
                >{\centering\arraybackslash}X}
                \toprule
                    \multicolumn{2}{c}{Bandwidths}  & Cluster (real) & Python prototype & \wrench simulator\\
                \midrule
                \multirow{2}{*}{Memory}      & read  & 6860 & 4812 & 4812\\
                                             & write & 2764 & 4812 & 4812\\
                \multirow{2}{*}{Local disk}  & read  & 510  & 465  & 465\\
                                             & write & 420  & 465  & 465\\
                \multirow{2}{*}{Remote disk} & read  & 515  & -    & 445\\
                                             & write & 375  & -    & 445\\
                \multicolumn{2}{l}{Network}  & 3000  & -    & 3000\\
                \bottomrule
                \end{tabularx}
                \caption{Bandwidth benchmarks (MBps) and simulator configurations.
                The bandwidths used in the simulations were the average of the measured read and write bandwidths.
                Network accesses were not simulated in the Python prototype.}
                \label{table:benchmark}
                \end{table}

        We used the real application in the fourth experiment
        (\textit{Exp~4}), which runs on a single cluster node 
        using a single local disk, to process data from
        participant 0027430 in the dataset of the Max Planck Institute for
        Human Cognitive and Brain Sciences available at
        \url{http://dx.doi.org/10.15387/fcp_indi.corr.mpg1}, leading to the
        parameters in Table~\ref{table:nighres_stats}.

        To calibrate the simulators, we benchmarked the
        memory, local disk, remote disk (NFS), and network bandwidths
        (Table~\ref{table:benchmark}). Since \simgrid, and thus \wrench, currently only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.

@@

    \begin{figure*}
        \centering
        \begin{subfigure}{\linewidth}
            \centering
               \includegraphics[width=\linewidth]{result/single/figures/single_errors_v2.pdf}
               \vspace*{-0.7cm}
               \caption{Relative simulation errors (absolute values)}
               \vspace*{0.5cm}
               \label{fig:single_error}
            \end{subfigure}
        \begin{subfigure}{\linewidth}
            \centering
            %    Gray shades represent task phases (read, compute and write).
            %    Lines represent memory usage along pipeline execution time.}
               \includegraphics[width=\linewidth]{result/single/figures/single_memprof_v2.pdf}
               \vspace*{-0.7cm}
               \caption{Memory profiles}
               \vspace*{0.5cm}
               \label{fig:single_memprof}
        \end{subfigure}
        \begin{subfigure}{\linewidth}
            \centering
               \includegraphics[width=1.05\linewidth]{result/single/figures/cached_files_v2.pdf}
               \caption{Page cache content after each I/O operation}
            %    \textcolor{red}{Update real results of 20GB}}
               \label{fig:single_cache}
        \end{subfigure}
        \caption{Single-threaded results (\textit{Exp 1})}
        \end{figure*}

@@

        \wrench simulation errors are substantially lower with 100~GB
        files than with 20~GB files, due to the fact that part of the
        100~GB file needs to be read and written to disk as it does not fit in cache. Disk models are
        identical in WRENCH and WRENCH-cache.

        Simulated memory profiles are highly consistent with the real ones
        (Figure~\ref{fig:single_memprof}). With 20~GB files, memory profiles almost exactly match the
        real ones, although for some reason dirty data seems to be flushing faster in real
        life than in simulation, a behavior also
        observed with 100~GB files. With 100~GB files, used memory reaches
        total memory during the first write, triggering dirty data
        flushing, and drops back to cached memory when application tasks
        release anonymous memory. Simulated cached memory is highly
        consistent with real values. In all cases, dirty data remains under the dirty ratio as
        expected. The Python prototype and \wrench-cache exhibit nearly
        identical memory profiles, which reinforces the confidence in our
        implementations.

        The content of the simulated memory cache is also highly consistent
        with reality (Figure~\ref{fig:single_cache}), showing that the
        previous observations were not incidental.

            \begin{figure*}
                \begin{subfigure}{\linewidth}
                    \centering
                    \includegraphics[width=0.8\linewidth]{result/multi/figures/multi_local.pdf}
                \end{subfigure}
                \caption{Concurrent results with 3~GB files (\textit{Exp 2})}
                \label{fig:multi_local}
            \end{figure*}
            \begin{figure*}
                \begin{subfigure}{\linewidth}
                    \centering
                    \includegraphics[width=0.8\linewidth]{result/multi/figures/multi_nfs.pdf}
                \end{subfigure}
                \caption{NFS results with 3~GB files (\textit{Exp 3})}
                \label{fig:multi_nfs}
                \end{figure*}
        \subsection{Concurrent applications (Exp 2)}

@@

@@

        \begin{figure}[b]
                \centering
                \includegraphics[width=0.95\linewidth]{result/nighres/figures/nighres_errors.pdf}
            \caption{Real application results (\textit{Exp 4})}
            \label{fig:nighres}
            \end{figure}

@@

        As is the case for \wrench, simulation time with \wrench-cache scales
        linearly with the number of concurrent applications
        (Figure~\ref{fig:multi_time}, p \textless $10^{-17}$). However, the page
        cache model substantially increases simulation time by
        application, as can be seen by comparing regression slopes in
        Figure~\ref{fig:multi_time}. Interestingly, \wrench-cache is faster with 
        NFS I/Os than with local I/Os, most likely due to the use of writethrough
        cache in NFS, which bypasses flushing operations.

        \begin{figure}
            \centering
            \includegraphics[width=0.95\linewidth]{result/multi/figures/simulation_time.pdf}
            \caption{Simulation time comparison. \wrench-cache scales
            linearly with the number of concurrent applications, albeit
            with a higher overhead than \wrench.}
            \label{fig:multi_time}
            \end{figure}

@@

\clearpage
\bibliographystyle{IEEEtran}
\bibliography{citation}
